---
title: "PredictingStroke"
author: "Jake McAferty"
date: "December 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(caret)
options("scipen"=100)
```
# Predicting Stroke

## Data
The data from this project can be found [on Kaggle](https://www.kaggle.com/asaumya/healthcare-dataset-stroke-data).  Test data does not contain stroke indicator so train file is split between test and train data sets.

### Load
```{r LoadData}
data <- read.csv("train_2v.csv")
#data.test <- read.csv("test_2v.csv")
```

### Raw Data Statistics

```{r SummaryStatistics; echo = FALSE}
summary.stats <- function(data){
    list(rowcount = nrow(data),
         nstrokes = data %>% filter(stroke == 1) %>% nrow() %>% as.numeric()
         )
}
summary.stats(data)
```

### Cleaning
* Gender of "Other" was removed (11 observations)
* BMI over 70 was removed (17 observations)
* Ages less than or equal to 1 were removed (530 observations)

```{r CleanData}
#Define Logical Variables
logicvars <- c("hypertension", "heart_disease", "stroke")

data.train <- data %>%
  filter(gender != "Other",
         bmi < 70,
         age > 1) %>%
  mutate_at(logicvars, as.logical)

#Print row counts of removed data
print("'Other' Gender")
data %>% filter(gender == "Other") %>% nrow()
print("BMI over 70")
data %>% filter(bmi >= 70) %>% nrow()
print("Age over 1")
data %>% filter(age <= 1) %>% nrow()
```

### Feature Engineering

* Bucket ages in 5 year increments
```{r Features}
#Add cusom features
data.train <- data.train %>%
  mutate(age_bucket = cut(age, seq(0, 85, by = 5)))
```
## Exploratory Analysis

```{r Exploration, echo=FALSE}
library(gridExtra)

summary(data.train)

hist.bmi <- data.train %>%
  ggplot(mapping = aes(x = bmi, fill = stroke)) +
  geom_histogram() +
  stat_density()

hist.age <- data.train %>%
  ggplot(mapping = aes(x = age_bucket, fill = stroke)) +
  geom_bar()

hist.gluc <- data.train %>%
  ggplot(mapping = aes(x = avg_glucose_level, fill = stroke)) +
  geom_histogram()

sp.agebmi <- data.train %>%
  ggplot(mapping = aes())

grid.arrange(hist.bmi, hist.gluc, ncol = 2)
hist.age

tbl.wrktype <-  table(data.train$stroke, data.train$work_type)

print("Stroke vs. Work Type")
tbl.wrktype
```
BMI is approximately normal around 25, glucose appears to be bimodal around 100 and 200.  Appears to be an uptick in strokes at higher ages.  

Several observations do not have smoker status, so this variable should be used with caution in any models.

Self employed individuals have a stroke rate 3 times higher than other working groups.

## Model
Two models will be created to predict if an individual has a 

1. Logistic Regression
2. Random forrest

### Model Setup
```{r ModelSetup}
set.seed(100)
model.vars <- c("age", 
                "avg_glucose_level", 
                "bmi", 
                "heart_disease",
                "hypertension")

model.target <- "stroke"
model.formula <- paste(c(model.target,
                         paste(model.vars, collapse = " + ")),
                       collapse = " ~ ") %>% as.formula()
model.formula

model.splitindices <- createDataPartition(data.train$stroke, p = .8, list = FALSE)

data.t <- data.train[model.splitindices,]
data.v <- data.train[-model.splitindices,]


print(nrow(samp.stroke)/nrow(data.t)) #Less than 2% of observations have a stroke

#Undersample training data
samp.count <- data.t %>% filter(stroke) %>% nrow()*2 #make a third of observations be true
samp.nostroke <- data.t %>% filter(!stroke)
samp.stroke <- data.t %>% filter(stroke)

samp.ind <- sample(x = seq(1, nrow(samp.nostroke)), size = samp.count)
data.t <- rbind(samp.nostroke[samp.ind,], samp.stroke)

print(nrow(samp.stroke)/nrow(data.t))
```

### Logistic Regression Model
```{r TreeModel}
set.seed(100)
model.glm <- glm(formula = model.formula,
                 family = binomial(link = "logit"),
                 data = data.t)

summary(model.glm)

model.glm.pred <- predict(model.glm, type = "response")
confusionMatrix(as.factor(model.glm.pred > .5), reference = factor(data.t$stroke))

model.glm.predv <- predict(model.glm, newdata = data.v,type = "response")
confusionMatrix(as.factor(model.glm.predv > .5), reference = factor(data.v$stroke))
```